### 2020-01-28 CalNotes.tex

In the following, we consider the filter with $\beta = 1$, i.e.,
\begin{equation}
	k(t,s) = \frac{1}{\delta} e^{-\frac{t-s}{\delta}},
\end{equation}
since in this case it holds
\begin{equation}
	\d Z^\epl_t = k(t,t) X^\epl_t \dd t + \int_0^t \partial_t k(t,s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{equation}
Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain a system of stochastic differential equations
\begin{equation}
\label{systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
Let $\rho^\epl = \rho^\epl(x,z)$ be the stationary distribution of the joint process $(X^\epl, Z^\epl)^\top$. \corr{How do we show that the couple $(X^\epl, Z^\epl)^\top$ is ergodic.} Then $\rho^\epl$ is the solution of the two-dimensional stationary Fokker--Planck equation
\begin{equation}
\begin{aligned}
	\label{eq:FPsystem}
	\sigma \partial^2_{xx} \rho^\epl(x,z) & + \left ( \alpha \cdot V''(x) + \frac{1}{\epl^2} p'' \left ( \frac{x}\epl \right ) + \frac{1}{\delta} \right ) \rho^\epl(x,z) \\
	& + \left ( \alpha \cdot V'(x) + \frac{1}\epl p' \left ( \frac{x}\epl \right ) \right ) \partial_x \rho^\epl(x,z) + \frac{1}{\delta}(z - x) \partial_z \rho^\epl(x,z) = 0.
\end{aligned}
\end{equation}

@@

\begin{lemma}\label{lem:FPMarginal} Let $\rho^\epl$ be the \corr{unique solution} of \eqref{eq:FPsystem} and let us write 
\begin{equation}\label{eq:densityDecomposition}
	\rho^\epl(x, z) = \phi^\epl(z)\psi^\epl(z)R^\epl(x,z),
\end{equation}
where $\phi^\epl$ and $\psi^\epl$ are the marginal densities, i.e., 
\begin{equation}
	\phi^\epl(x) = \int_{\R} \rho^\epl(x,z) \dd z, \quad  \psi^\epl(z) = \int_{\R} \rho^\epl(x,z) \dd x.
\end{equation}
Then, it holds
\begin{equation}\label{eq:marginalX}
	\phi^\epl(x) = \frac{1}{C_{\phi^\epl}} \exp\left(-\frac{1}{\sigma} \alpha \cdot V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right),
\end{equation}
where
\begin{equation}
	C_{\phi^\epl} = \int_{\R} \exp\left(-\frac{1}{\sigma} \alpha \cdot V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right) \dd x.
\end{equation}
Moreover, it holds
\begin{equation}
	\sigma \delta \int_{\R} \int_{\R} V'(z) \phi^\epl(x) \psi^\epl(z) \partial_x R^\epl(x,z) \dd x \dd z = \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)].
\end{equation}
\end{lemma}

>>>>

In the following, we consider the filter with $\beta = 1$, i.e.,
\begin{equation}
	k(t,s) = \frac{1}{\delta} e^{-\frac{t-s}{\delta}},
\end{equation}
since in this case it holds
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t \partial_t k(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{equation}

We will treat analytically two different cases. First, the case where the state $X_t^\epl$ is a vector of $\R^d$ and we wish to find the

Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain a system of stochastic differential equations
\begin{equation}
\label{systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
Let $\rho^\epl = \rho^\epl(x,z)$ be the stationary distribution of the joint process $(X^\epl, Z^\epl)^\top$. \corr{How do we show that the couple $(X^\epl, Z^\epl)^\top$ is ergodic.} Then $\rho^\epl$ is the solution of the two-dimensional stationary Fokker--Planck equation
\begin{equation}
\begin{aligned}
	\label{eq:FPsystem}
	\sigma \partial^2_{xx} \rho^\epl(x,z) & + \left ( \alpha \cdot V''(x) + \frac{1}{\epl^2} p'' \left ( \frac{x}\epl \right ) + \frac{1}{\delta} \right ) \rho^\epl(x,z) \\
	& + \left ( \alpha \cdot V'(x) + \frac{1}\epl p' \left ( \frac{x}\epl \right ) \right ) \partial_x \rho^\epl(x,z) + \frac{1}{\delta}(z - x) \partial_z \rho^\epl(x,z) = 0.
\end{aligned}
\end{equation}

@@

\begin{lemma}\label{lem:FPMarginal} Let $\rho^\epl$ be the \corr{unique solution} of \eqref{eq:FPsystem} and let us write 
\begin{equation}\label{eq:densityDecomposition}
	\rho^\epl(x, z) = \phi^\epl(z)\psi^\epl(z)R^\epl(x,z),
\end{equation}
where $\phi^\epl$ and $\psi^\epl$ are the marginal densities, i.e., 
\begin{equation}
	\phi^\epl(x) = \int_{\R} \rho^\epl(x,z) \dd z, \quad  \psi^\epl(z) = \int_{\R} \rho^\epl(x,z) \dd x.
\end{equation}
Then, it holds
\begin{equation}\label{eq:marginalX}
	\phi^\epl(x) = \frac{1}{C_{\phi^\epl}} \exp\left(-\frac{1}{\sigma} \alpha \cdot V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right),
\end{equation}
where
\begin{equation}
	C_{\phi^\epl} = \int_{\R} \exp\left(-\frac{1}{\sigma} \alpha \cdot V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right) \dd x.
\end{equation}
Moreover, it holds
\begin{equation}
	\sigma \delta \int_{\R} \int_{\R} \phi^\epl(x) \psi^\epl(z) \nabla_z V(z) \cdot \nabla_x R^\epl(x,z) \dd x \dd z = \E^{\rho^\epl}[((X^\epl)^2 - (Z^\epl)^2)V''(Z^\epl)].
\end{equation}
\end{lemma}
